Aspects of RDF Data Constraints in the Social, Behavioural and Economic Sciences

### 2015-04-30 paper/paper7/paper7.tex
4

%{Simple constraint types} is the set of constraint types whose constraints can be easily defined without much effort in addition to $\mathcal{CT}_{B}$ constraints.
For data properties, it may be desirable to restrict that values of predefined languages must be present for determined number of times (\emph{R-48/49: language tag cardinality}):
(1) It is checked if literal language tags are set. Some controlled vocabularies, e.g., contain literals in natural language, but without information what language has actually been used (\emph{warning}). 
(2) Language tags must conform to language standards (\emph{error}). 
(3) Some thesaurus concepts are labeled in only one, others in multiple languages. 
It may be desirable to have each concept labeled in each of the languages that are also used on the other concepts,
as language coverage incompleteness for some concepts may indicate shortcomings of thesauri (\emph{informational})
\cite{MaderHaslhoferIsaac2012}.
%\textbf{Validation and Reasoning.}
%\emph{Default values} (\emph{R-31, R-38}) for objects/literals of given properties are inferred automatically when properties are not present in the data.
%The value \emph{true} for the property {\em disco:isPublic} indicates that a {\em disco:LogicalDataSet} can be accessed by anyone.
%Per default, however, access to data sets should be restricted (\emph{false}) ($\mathcal{SL}_{0}$).
%Many properties are not necessarily required but \emph{recommended} within a particular context (\emph{R-72}).
%The property {\em skos:notation}, e.g., is not mandatory for {\em disco:Variable}s, but recommended to represent variable names ($\mathcal{SL}_{0}$).
Percentage values are only valid when they are within the literal range of 0 and 100 (\emph{R-45: literal ranges}; $\mathcal{SL}_{2}$)
which is checked for \emph{disco:percentage} standing for the number of cases of a given code in relation to the total number of cases for a particular variable.

@@

%$\mathcal{CT}_{C}$ denotes the set of constraint types for which the definition of constraints is rather complex and cannot be derived from vocabulary definitions.
\emph{Data model consistency} constraints ensure the integrity of
the data according to the intended semantics of vocabularies.
Every \emph{qb:Observation}, e.g., must have a value for each dimension
declared in its \emph{qb:DataStructureDefinition} ($\mathcal{SL}_{2}$)
and no two \emph{qb:Observations} in the same \emph{qb:DataSet}
can have the same value for all dimensions (\emph{warning}).
If a \emph{qb:DataSet} \emph{D} has a \emph{qb:Slice} \emph{S}, and \emph{S} has an
\emph{qb:Observation} \emph{O}, then the \emph{qb:DataSet} corresponding to \emph{O} must be \emph{D} (\emph{warning}).
%Relative frequencies of variable codes are calculated correctly, if the cumulative percentage (\emph{disco:cumulativePercentage}) of a given code exactly matches the cumulative percentage of the previous code
%plus the percentage value (\emph{disco:percentage}) of the current code ($\mathcal{SL}_{2}$).
%{\em Mathematical Operations} (\emph{R-41, R-42}; e.g. date calculations and statistical computations like average, mean, and sum) are performed to ensure the integrity of data models.
%The sum of percentage values of all variable codes, e.g., must exactly be 100 ($\mathcal{SL}_{2}$)
%and the minimum absolute frequency of all variable codes do not have to be greater than the maximum ($\mathcal{SL}_{2}$).

%\textbf{Hierarchies and Ordering.}
%\textbf{Structure and Ordering.}
%For assessing the quality of \emph{SKOS} vocabularies, we concentrate on graph-based structures and apply graph- and network-analysis techniques (\emph{structure} constraints) like 
%(1) a thesaurus should provide entry points (top concepts) to the data to provide efficient access and guidance for human users,
%(2) concepts, internal to the tree, should not be indicated as top concepts, and
%(3) a thesaurus should not contain many orphan concepts 
%(concepts without any associative or hierarchical relations) lacking valuable context information for retrieval.
%, as, e.g., no hierarchical query expansion can be performed on search terms to find documents with more general content.) \cite{MaderHaslhoferIsaac2012}.

%A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
%To compare variables 
%(1) their code lists must be structured properly and
%(2) their code lists must either be identical or at least similar.
%If a researcher only wants to get a first overview over comparable variables (use case 1), 
%covering the first constraint may be sufficient.
%Thus, the severity level of the first constraint is stronger ($\mathcal{SL}_{2}$) than the one for the second constraint ($\mathcal{SL}_{0}$).
%If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the second constraint is getting more serious
%and the user may raise its severity level.
%\textbf{Comparison.}
%A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
%To compare variables, 
%(1) variables and (2) variable definitions must be present,
%(3) code lists must be structured properly,
%(4) for each code an associated category must be specified, and
%(5) code lists must either be identical or at least similar.
%If a researcher wants to get a first overview over comparable variables (use case 1), 
%covering the first three constraints may be sufficient.
%Thus, the severity level of the first three constraints is stronger ($\mathcal{SL}_{2}$) than for the last two constraints ($\mathcal{SL}_{1}$ and $\mathcal{SL}_{0}$).
%If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the remaining two constraints is getting more serious.

%\textbf{Membership in Controlled Vocabularies.}
In many cases, resources must be \emph{members of controlled vocabularies} (\emph{R-32}).
If a dimension property, e.g., has a \emph{qb:codeList},
then the value of the dimension property on every \emph{qb:Observation} must be in the code list ($\mathcal{SL}_{2}$).
Summary statistics types like minimum, maximum, and arithmetic mean are maintained within a controlled vocabulary.  
Thus, summary statistics can only have \emph{disco:summaryStatisticType} relationships to \emph{skos:Concept}s which must be members of the controlled vocabulary \emph{ddicv:SummaryStatisticType}, a \emph{skos:ConceptScheme} ($\mathcal{SL}_{2}$).
Objects/literals can be declared to be ordered for given properties (\emph{R-121/217: ordering}).
Variables, questions, and codes, e.g., are typically organized in a particular order. 
If codes (\emph{skos:Concept}) should be ordered, they must be members (\emph{skos:memberList}) in an ordered collection (\emph{skos:OrderedCollection}), the variable's code list ($\mathcal{SL}_{0}$).

%\textbf{Constraints on Properties.}
It is useful to declare properties to be \emph{conditional} (\emph{R-71}), i.e., if particular properties exist (or do not exist), then other properties must also be present (or absent).
To get an overview over a series/study either an abstract, a title, an alternative title, or links to external descriptions should be provided. 
If an abstract and an external description are absent, however,  
a title or an alternative title should be given (\emph{warning}).
In case a variable is represented in form of a code list, codes may be associated with categories, i.e., human-readable labels ($\mathcal{SL}_{0}$).
The variable \emph{Education at pre-school}, e.g., is represented as ordered code list without any categories.
%If a {\em skos:Concept} represents a code (having {\em skos:notation} and {\em skos:prefLabel} properties), 
%then the property {\em disco:isValid} has to be stated indicating if the code stands for valid (\emph{true}) or missing (\emph{false}) cases ($\mathcal{SL}_{2}$).
\emph{Context-specific exclusive or of property groups} (\emph{R-11}) constraints
restrict individuals of given classes to have properties defined within exactly one of multiple property groups.
\emph{skos:Concept}s can have either \emph{skos:definition} (when interpreted as theoretical concepts) or \emph{skos:notation} and \emph{skos:prefLabel} properties (when interpreted as codes), but not both ($\mathcal{SL}_{2}$).

>>>>

%{Simple constraint types} is the set of constraint types whose constraints can be easily defined without much effort in addition to $\mathcal{CT}_{B}$ constraints.
For data properties, it may be desirable to restrict that values of predefined languages must be present for determined number of times (\emph{R-48/49: language tag cardinality}):
(1) It is checked if literal language tags are set. Some controlled vocabularies, e.g., contain literals in natural language, but without information what language has actually been used (\emph{warning}). 
(2) Language tags must conform to language standards (\emph{error}). 
(3) Some thesaurus concepts are labeled in only one, others in multiple languages. 
It may be desirable to have each concept labeled in each of the languages that are also used on the other concepts,
as language coverage incompleteness for some concepts may indicate shortcomings of thesauri (\emph{informational})
\cite{MaderHaslhoferIsaac2012}.
%\textbf{Validation and Reasoning.}
%\emph{Default values} (\emph{R-31, R-38}) for objects/literals of given properties are inferred automatically when properties are not present in the data.
%The value \emph{true} for the property {\em disco:isPublic} indicates that a {\em disco:LogicalDataSet} can be accessed by anyone.
%Per default, however, access to data sets should be restricted (\emph{false}) ($\mathcal{SL}_{0}$).
%Many properties are not necessarily required but \emph{recommended} within a particular context (\emph{R-72}).
%The property {\em skos:notation}, e.g., is not mandatory for {\em disco:Variable}s, but recommended to represent variable names ($\mathcal{SL}_{0}$).
Percentage values are only valid when they are within the literal range of 0 and 100 (\emph{R-45: literal ranges}; \emph{error})
which is checked for \emph{disco:percentage} standing for the number of cases of a given code in relation to the total number of cases for a particular variable.

@@

%$\mathcal{CT}_{C}$ denotes the set of constraint types for which the definition of constraints is rather complex and cannot be derived from vocabulary definitions.
\emph{Data model consistency} constraints ensure the integrity of
the data according to the intended semantics of vocabularies.
Every \emph{qb:Observation}, e.g., must have a value for each dimension
declared in its \emph{qb:DataStructureDefinition} (\emph{error})
and no two \emph{qb:Observations} in the same \emph{qb:DataSet}
can have the same value for all dimensions (\emph{warning}).
If a \emph{qb:DataSet} \emph{D} has a \emph{qb:Slice} \emph{S}, and \emph{S} has an
\emph{qb:Observation} \emph{O}, then the \emph{qb:DataSet} corresponding to \emph{O} must be \emph{D} (\emph{warning}).
%Relative frequencies of variable codes are calculated correctly, if the cumulative percentage (\emph{disco:cumulativePercentage}) of a given code exactly matches the cumulative percentage of the previous code
%plus the percentage value (\emph{disco:percentage}) of the current code ($\mathcal{SL}_{2}$).
%The sum of percentage values of all variable codes, e.g., must exactly be 100 ($\mathcal{SL}_{2}$)
%and the minimum absolute frequency of all variable codes do not have to be greater than the maximum ($\mathcal{SL}_{2}$).
%\textbf{Structure and Ordering.}
%For assessing the quality of \emph{SKOS} vocabularies, we concentrate on graph-based structures and apply graph- and network-analysis techniques (\emph{structure} constraints) like 
%(1) a thesaurus should provide entry points (top concepts) to the data to provide efficient access and guidance for human users,
%(3) a thesaurus should not contain many orphan concepts 
%(concepts without any associative or hierarchical relations) lacking valuable context information for retrieval.
%, as, e.g., no hierarchical query expansion can be performed on search terms to find documents with more general content.) \cite{MaderHaslhoferIsaac2012}. 
%A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
%To compare variables 
%(1) their code lists must be structured properly and
%(2) their code lists must either be identical or at least similar.
%If a researcher only wants to get a first overview over comparable variables (use case 1), 
%covering the first constraint may be sufficient.
%Thus, the severity level of the first constraint is stronger ($\mathcal{SL}_{2}$) than the one for the second constraint ($\mathcal{SL}_{0}$).
%If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the second constraint is getting more serious
%and the user may raise its severity level.
%\textbf{Comparison.}
%A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
%(4) for each code an associated category must be specified, and
%covering the first three constraints may be sufficient.
%Thus, the severity level of the first three constraints is stronger ($\mathcal{SL}_{2}$) than for the last two constraints ($\mathcal{SL}_{1}$ and $\mathcal{SL}_{0}$).
In many cases, resources must be \emph{members of controlled vocabularies} (\emph{R-32}).
If a dimension property, e.g., has a \emph{qb:codeList},
then the value of the dimension property on every \emph{qb:Observation} must be in the code list (\emph{error}).
%Thus, summary statistics can only have \emph{disco:summaryStatisticType} relationships to \emph{skos:Concept}s which must be members of the controlled vocabulary \emph{ddicv:SummaryStatisticType}, a \emph{skos:ConceptScheme} ($\mathcal{SL}_{2}$).
Objects/literals can be declared to be ordered for given properties (\emph{R-121/217: ordering}).
Variables, questions, and codes, e.g., are typically organized in a particular order. 
If codes (\emph{skos:Concept}) should be ordered, they must be members (\emph{skos:memberList}) in an ordered collection (\emph{skos:OrderedCollection}), the variable's code list (\emph{informational}).

%\textbf{Constraints on Properties.}
It is useful to declare properties to be \emph{conditional} (\emph{R-71}), i.e., if particular properties exist (or do not exist), then other properties must also be present (or absent).
To get an overview over a series/study either an abstract, a title, an alternative title, or links to external descriptions should be provided. 
If an abstract and an external description are absent, however,  
a title or an alternative title should be given (\emph{warning}).
In case a variable is represented in form of a code list, codes may be associated with categories, i.e., human-readable labels (\emph{informational}).
The variable \emph{Education at pre-school}, e.g., is represented as ordered code list without any categories.
%If a {\em skos:Concept} represents a code (having {\em skos:notation} and {\em skos:prefLabel} properties), 
%then the property {\em disco:isValid} has to be stated indicating if the code stands for valid (\emph{true}) or missing (\emph{false}) cases ($\mathcal{SL}_{2}$).
\emph{Context-specific exclusive or of property groups} (\emph{R-11}) constraints
restrict individuals of given classes to have properties defined within exactly one of multiple property groups.
\emph{skos:Concept}s can have either \emph{skos:definition} (when interpreted as theoretical concepts) or \emph{skos:notation} and \emph{skos:prefLabel} properties (when interpreted as codes), but not both (\emph{error}).
